---
title: "Week2 Rmarkdown"
author: "Tianheng Z"
date: "06/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = list(CRAN="http://cran.rstudio.com/"))
```

## Include and Install the library
```{r,echo=TRUE,eval=TRUE,results='hide'}
install.packages("tidyverse")
library(tidyverse)
install.packages("Stat2Data")
library(Stat2Data)
install.packages("ggpubr")
library(ggpubr)
```

# Load the data
```{r}
data("Hawks")
hawksSmall<-drop_na(select(Hawks,Age,Day,Month,Year,
                           CaptureTime,Species,Wing,Weight,Tail))
```
# 1.Visualization
## 1.1 Type of variables
### Check the dim and head of data
```{r}
dim(hawksSmall)
head(hawksSmall,5)
names(hawksSmall)
```
### Month-> Categorical, Species->Categorical, Age->discrete, Wing-> Discrete, Weight->Discrete

## 1.2 What's wrong with the plot
### There are too many atttibutes in the plot, making it not easy to understatnd.

## 1.3 Generate a histogram
```{r}
#Bimodal data
ggplot(data = hawksSmall,aes(x=Weight)) + geom_histogram(binwidth = 100)
```

## 1.4 Generate a density plot
```{r}
density_half <-ggplot(data = hawksSmall,aes(x = Tail)) + geom_density(adjust =
        0.5) + xlab("Tail(mm)") + ggtitle("adjust = 0.5")

density_one <- ggplot(data = hawksSmall,aes(x = Tail)) + geom_density(adjust = 
          1) + xlab("Tail(mm)") +ggtitle("adjust = 1")

ggarrange(density_half,density_one,nrow = 2,ncol = 1)
```
```{r}
ggplot(data = hawksSmall,aes(x = Tail,color = Species)) + geom_density() + xlab("Tail(mm)")
```
```{r}
ggplot(data = hawksSmall,aes(x = Tail,y = Species,fill = Species)) + geom_violin() + xlab("Tail(mm)")
```

## 1.5 Scatter plots
```{r}
ggplot(data = hawksSmall,aes(x = Tail, y = Weight,color = Species,shape = Species))+
  geom_point() + xlab("Tail(mm)") + ylab("Weight(gm)")
```

## 1.6 Trend Lines and facet wraps
```{r}
ggplot(data = hawksSmall,aes(x = Tail, y =Weight,color = Species)) + geom_point()+ geom_smooth(method = "lm")+facet_wrap(~Species) +xlab("Tail(mm)") + ylab("Weight(mm)")
```

# 2. Data Wrangling
## 2.1 Select and filter functions
```{r}
hSF <- Hawks %>% filter(Species == "RT", Weight >=1000) %>% select(Wing,Weight,Tail) 
dim(hSF)
```

## 2.2 The arrange function

```{r}
head(hSF %>% arrange(Wing),5)

```

## 2.3 Join and rename funtions
```{r}
inter <- data.frame(species_code = c("CH","RT","SS"),
                    Species_name_full = c("Cooper's","Red-tailed","Sharp-shinned"))

hawksFullname <-left_join(Hawks,rename(inter,Species = species_code),by ="Species" )
hawksFullname %>% select(Species_name_full,Wing,Weight) %>% rename(Species = Species_name_full) %>% head(7)
```

## 2.4 The mutate function

```{r}
hawksWithBMI <- Hawks %>% mutate(bird_BMI = 1000 * Weight/(Wing^2) ) %>%
  select(Species,bird_BMI) %>% arrange(desc(bird_BMI)) 
head(hawksWithBMI,8)
```
```{r}
hawksWithBMI %>% filter(bird_BMI < 100) %>% 
  ggplot(aes(y=Species,x = bird_BMI,fill = Species))+
  geom_violin()
```

## 2.5 Summarize and group-by functions
```{r}
hawksFullname %>% 
  group_by(Species) %>% 
  summarise(num_rows = n(),mn_wing = mean(Wing,na.rm = 1),md_wing = median(Wing,na.rm = 1),
            t_mn_wing = mean(Wing,trim=0.1,na.rm = 1),
            tail_wing_ratio = mean(Wing/Tail,na.rm = 1))
```

```{r}
hawksFullname %>% select(Wing,Weight,Culmen,Hallux,Tail,StandardTail,Tarsus,Crop,Species_name_full) %>% rename(Species = Species_name_full) %>%group_by(Species) %>%summarise(across(everything(),~sum(is.na(.x))))
```


# 3. Exploratory data analysis

## 3.1 Combining location estimators with the summarise function
```{r}
Hawks %>% summarise(Wing_mean = mean(Wing,na.rm = 1),
                   Wing_t_mean = mean(Wing,trim = 0.1,na.rm = 1),
                   Wing_med = median(Wing,na.rm = 1),
                   Weight_mean = mean(Weight,na.rm = 1),
                   Weight_t_mean = mean(Weight,na.rm = 1,trim = 0.1),
                   Weight_med = median(Weight,na.rm = 1)
                   )

```
```{r}
Hawks %>% group_by(Species) %>% summarise(Wing_mean = mean(Wing,na.rm = 1),
                   Wing_t_mean = mean(Wing,trim = 0.1,na.rm = 1),
                   Wing_med = median(Wing,na.rm = 1),
                   Weight_mean = mean(Weight,na.rm = 1),
                   Weight_t_mean = mean(Weight,na.rm = 1,trim = 0.1),
                   Weight_med = median(Weight,na.rm = 1)
                   )
```

## 3.2 Location and dispersion estimatiors under linear transformation
```{r}
a = c(1,2,3,4,5,6,7,8,9,10)
mean(a)
var(a)
```

```{r}
mean(5 *a + 10)
var(5 *a +10)
```

## 3.3 Robustness of location estimators
```{r}
hal<-Hawks$Hallux     # Extract the vector of hallux lengths
hal<-hal[!is.na(hal)] # Remove any nans
outlier_val<-100
num_outliers<-10
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
mean(hal)
mean(corrupted_hal)
```
```{r}
num_outliers_vect<-seq(0,1000)
means_vect<-c()
for(num_outliers in num_outliers_vect){
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
means_vect<-c(means_vect,mean(corrupted_hal))
}
```


```{r}
num_outliers_vect<-seq(0,1000)
medians_vect <- c()
for(num_outliers in num_outliers_vect){
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
medians_vect<-c(medians_vect,median(corrupted_hal))
}
```


```{r}
num_outliers_vect<-seq(0,1000)
t_means_vect <- c()
for(num_outliers in num_outliers_vect){
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
t_means_vect<-c(t_means_vect,mean(corrupted_hal,trim = 0.1))
}
```

```{r}
df_means_medians<-data.frame(num_outliers=num_outliers_vect,
mean=means_vect,t_mean=t_means_vect,
median=medians_vect)



```

```{r}
df_means_medians%>%
pivot_longer(!num_outliers, names_to = "Estimator", values_to = "Value")%>%
ggplot(aes(x=num_outliers,color=Estimator,
linetype=Estimator,y=Value))+
geom_line()+xlab("Number of outliers")

```

## 3.4 Box plots and outliers
```{r}
Hawks %>% group_by(Species) %>% ggplot(aes(x = Species,y = Weight)) + geom_boxplot()
```
```{r}
num_outliers <-function(sample)
{
  
  out_num <- 0
  q25 <- quantile(sample,prob = .25)
  q75 <- quantile(sample,probs = .75)
  sec <- q75-q25
  for(num in sample)
  {
    if(num <q25-1.5*sec | num > q75 + 1.5*sec)
    {
      out_num = out_num +1
    }
  }
  out_num
}

Hawks %>%filter(!is.na(Weight)) %>% group_by(Species) %>% summarise(outlier_weights = num_outliers(Weight))
```
## 3.5 Covariance and Correlation under linear transformation
```{r}
set.seed(10000)
X = rnorm(100,5,5)
Y = rnorm(100,10,10)
cov(X,Y)
cov(X,Y)/sd(X)/sd(Y)
```
```{r}
new_X = 5*X + 10
new_Y = 2*Y + 4
cov(new_X,new_Y)
cov(new_X,new_Y)/sd(new_X)/sd(new_Y)
```



